Temporal Split

### 2023-01-30 workflow/notebooks/temporal-split.Rmd
compare OTU vs genus

@@

### vs 100x train/test splits

```{r perf_100x}
perf_dat_100 <- data.table::fread(here('results', 
                                      'performance_results_aggregated.csv')
                                 ) %>% 
  pivot_longer(c('cv_metric_AUC', "AUC"), 
               names_to = 'term', values_to = 'estimate') %>% 
  mutate(term = case_when(term == 'cv_metric_AUC' ~ 'train AUROC',
                          term == 'AUC' ~ 'test AUROC',
                          TRUE ~ term)
         ) 
perf_100_plot <- perf_dat_100 %>% 
  ggplot(aes(x = estimate, y = outcome, color = term)) +
  geom_vline(xintercept = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  xlim(0,1) +
  facet_wrap('dataset', ncol = 1) +
  coord_flip() +
  labs(title = '100x train/test splits',
       caption = 'Box: interquartile range. Whisker: min & max') +
  theme_sovacool() +
  theme(legend.position = 'top',
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.caption = element_text(hjust = 0)
  )
perf_100_plot
```
```{r perf_temporal_vs_100, fig.width=10, fig.height=5}
cowplot::plot_grid(perf_temp_plot, perf_100_plot, nrow = 1)
```

>>>>

### temporal split

@@

### 100x train/test splits
```{r perf_100x}
perf_dat_100 <- data.table::fread(here('results', 
                                      'performance_results_aggregated.csv')
                                 ) %>% 
  pivot_longer(c('cv_metric_AUC', "AUC"), 
               names_to = '.term', values_to = '.estimate') %>% 
  mutate(.term = case_when(.term == 'cv_metric_AUC' ~ 'train_AUROC',
                          .term == 'AUC' ~ 'test_AUROC',
                          TRUE ~ .term)
         ) %>% 
  select(method, seed, outcome, taxlevel, metric, dataset, trainfrac, .term, .estimate) %>% 
  filter(!is.na(.estimate))
```
```{r perf_100_tbl}
perf_dat_100 %>% 
  filter(method == 'rf', trainfrac == 0.8, taxlevel == 'OTU') %>% 
  group_by(dataset, outcome, .term) %>% 
  summarize(median_est = round(median(.estimate),3)) %>% 
  pivot_wider(names_from = dataset, values_from = median_est, names_prefix = 'median_perf_') %>% 
  arrange(desc(outcome), desc(.term)) %>% 
  kable()
perf_dat_100 %>% 
  filter(method == 'glmnet', trainfrac == 0.8, taxlevel == 'OTU') %>% 
  group_by(dataset, outcome, .term) %>% 
  summarize(median_est = round(median(.estimate),3)) %>% 
  pivot_wider(names_from = dataset, values_from = median_est, names_prefix = 'median_perf_') %>% 
  arrange(desc(outcome), desc(.term)) %>% 
  kable()
```

```{r perf_100_glmnet}
perf_100_glmnet_plot <- perf_dat_100 %>% 
  filter(method == 'glmnet', trainfrac == 0.8, taxlevel == 'OTU') %>% 
  ggplot(aes(x = .estimate, y = outcome, color = .term)) +
  geom_vline(xintercept = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  xlim(0,1) +
  facet_wrap('dataset', ncol = 1) +
  coord_flip() +
  labs(title = '100x train/test splits - glmnet',
       caption = 'Box: interquartile range. Whisker: min & max') +
  theme_sovacool() +
  theme(legend.position = 'top',
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.caption = element_text(hjust = 0)
  )
perf_100_glmnet_plot
```
```{r, perf_100_rf}
perf_100_rf_plot <- perf_dat_100 %>% 
  filter(method == 'rf', trainfrac == 0.8, taxlevel == 'OTU') %>% 
  ggplot(aes(x = .estimate, y = outcome, color = .term)) +
  geom_vline(xintercept = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  xlim(0,1) +
  facet_wrap('dataset', ncol = 2) +
  labs(title = '100x train/test splits - rf',
       caption = 'Box: interquartile range. Whisker: min & max') +
  theme_sovacool() +
  theme(legend.position = 'top',
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.caption = element_text(hjust = 0)
  )
perf_100_rf_plot
```

#### 100 splits - OTU vs genus

```{r perf_100_otu_vs_genus}
perf_dat_100 %>% 
  filter(method == 'glmnet', trainfrac == 0.8) %>% 
  ggplot(aes(x = .estimate, y = outcome, color = .term)) +
  geom_vline(xintercept = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  xlim(0,1) +
  facet_wrap(c('dataset', 'taxlevel')) +
  coord_flip() +
  labs(title = '100x train/test splits - glmnet',
       caption = 'Box: interquartile range. Whisker: min & max') +
  theme_sovacool() +
  theme(legend.position = 'top',
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.caption = element_text(hjust = 0)
  )
```

### temporal split vs 100x train/test splits

```{r perf_temporal_vs_100, fig.width=10, fig.height=5}
cowplot::plot_grid(perf_temp_plot, perf_100_glmnet_plot, nrow = 1)
```
